---
title: "432 Project B"
author: "Jacob James Rich & Steven Mayher"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: show
    self_contained: true
    number_sections: true
    df_print: paged
---

# Setup {-}

```{r setup, message = FALSE}
library(knitr)
library(rmdformats)

## Global options

opts_chunk$set(comment=NA)
opts_knit$set(width=75)
```

## Package Loading {-}

```{r load_packages, message = FALSE}
library(here) #here()
library(doBy) #summaryBy()
library(magrittr)
library(janitor)
library(conflicted) # conflict_prefer()

library(rms) # rcs(), includes Hmisc package
library(MASS) #polr()
library(nnet)
library(skimr)

#install.packages("countreg", repos="http://R-Forge.R-project.org")
library(countreg)

library(broom) # tidy() 
library(car) # outlierTest() 
library(GGally) # ggpairs()
library(mosaic)
library(patchwork)
library(naniar) # n_miss() 
library(ipumsr) # IPUMS data

library(tidyverse)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("sum", "base")

# Program "not in" operator
'%!in%' <- function(x,y)!('%in%'(x,y))
```

## Sources

-Jacob Kaplan's Concatenated Files: Uniform Crime Reporting (UCR) Program [Data](https://www.openicpsr.org/openicpsr/project/102263/version/V14/view?path=/openicpsr/102263/fcr:versions/V14/ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds.zip&type=file): Arrests by Age, Sex, and Race, 1974-2020

- Florida Uniform Crime Report (UCR) [Data](https://www.fdle.state.fl.us/FSAC/CJAB-Home/Uniform-Crime-Report/Data-Archives/UCR-Arrest-Data)

https://www.fdle.state.fl.us/FSAC/Data-Statistics/UCR-Offense-Data.aspx

# Import and Clean Data

## Import 1980 & 1990 FBI Data

```{r}
# Download all files
fbi_1990  <- readRDS(here("ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds", "ucr_arrests_monthly_all_crimes_race_sex_1990.rds")) %>%
  select(year, population, fips_state_county_code, sale_drug_total_tot_white, sale_drug_total_tot_black, sale_drug_total_tot_arrests)

fbi_1980 <- readRDS(here("ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds", "ucr_arrests_monthly_all_crimes_race_sex_1980.rds")) %>%
  select(year, population, fips_state_county_code, sale_drug_total_tot_white, sale_drug_total_tot_black, sale_drug_total_tot_arrests)

fbi_1980_1990 <- rbind(data = fbi_1990, data = fbi_1980)

# Remove `fbi_1990` and `fbi_1990` to save memory
rm(fbi_1990, fbi_1980)
```

## Import 1980 & 1990 IPUMS Data

```{r}
ddi <- read_ipums_ddi("usa_00005.xml")

ipums <- read_ipums_micro(ddi)

ipums
```

## Build Panels

### FBI county level

```{r}
# Collapse by year and fips_state_county_code
fbi_county_years <- summaryBy(population + sale_drug_total_tot_white + sale_drug_total_tot_black + sale_drug_total_tot_arrests ~ year + fips_state_county_code, data=fbi_1980_1990, FUN = sum, keep.names = TRUE)

# Outcome variable calculations / remove puerto rico, guam, american samoa, virgin islands, florida / remove nonmatched counties
fbi_county_years <- fbi_county_years %>% 
  mutate(arrest_rate = sale_drug_total_tot_arrests / population * 100000,
         proportion_black = sale_drug_total_tot_black / sale_drug_total_tot_arrests,
         proportion_white = sale_drug_total_tot_white / sale_drug_total_tot_arrests,
         fips_state = substr(fips_state_county_code, 1, 2)) %>%
  filter(fips_state_county_code %!in% c(02188, 02100, 02060, 04012, 02060, 02070, 02013, 	04012),
         fips_state %!in% c(72, 66, 60, 78, 12))

fbi_1980_1990 <- fbi_county_years

#Remove redundant `fbi_county_years` to save memory
rm(fbi_county_years)
```

### IPUMS county level

```{r}
# Calculate individual weights
ipums <- ipums %>% 
  mutate(sex_base = if_else(SEX %in% c(1,2), PERWT, 0),
         male = if_else(SEX %in% c(1), PERWT, 0),
         female = if_else(SEX %in% c(2), PERWT, 0),
         income_base = if_else(INCWAGE %!in% c(999999, 999998), PERWT, 0),
         income = if_else(INCWAGE %!in% c(999999, 999998), INCWAGE*PERWT, 0),
         poverty_base = if_else(POVERTY %!in% c(000), PERWT, 0),
         poverty = if_else(POVERTY < 100, if_else(POVERTY == 0, 0, PERWT),0),
         age_base = if_else(AGE < 91, PERWT, 0),
         age = if_else(AGE < 91, AGE*PERWT, 0),
         race_base = if_else(RACE < 8, PERWT, 0),
         black = if_else(RACE %in% c(2), PERWT, 0),
         white = if_else(RACE %in% c(1), PERWT, 0),
         hispanic_base = if_else(HISPAN %!in% c(9), PERWT, 0),
         hispanic = if_else(HISPAN %in% c(1, 2, 3, 4), PERWT, 0),
         fips_state_county_code = if_else(nchar(COUNTYFIP) == 2, paste(STATEFIP, COUNTYFIP, sep = '0'),
                                          if_else(nchar(COUNTYFIP) == 1, paste(STATEFIP, COUNTYFIP, sep = '00'), paste(STATEFIP, COUNTYFIP, sep = '')))) %>% 
  rename(year = YEAR)

# Collapse into counties
ipums_1980_1990 <- summaryBy(sex_base + male + female + income_base + income + poverty_base + poverty + age_base + age + race_base + black + white + hispanic_base + hispanic ~ year + fips_state_county_code, data=ipums, FUN = sum, keep.names = TRUE)

# Remove ipums data to save memory
rm(ipums)

# Create IPUMS predictor variables
ipums_1980_1990 <- ipums_1980_1990 %>% 
  mutate(male = male / sex_base,
         female = female / sex_base,
         income = income / income_base,
         poverty = poverty / poverty_base,
         age = age / age_base,
         black = black / race_base,
         white = white / race_base,
         hispanic = hispanic / hispanic_base) %>% 
  select(year, fips_state_county_code, male, female, income, poverty, age, black, white, hispanic)
```

## Separate and combine panels

```{r}
# Separate `ipums_1980_1990` into individual years
ipums_1980 <- ipums_1980_1990 %>% 
  filter(year == 1980) %>% 
  mutate(male1980 = male,
         female1980 = female,
         income1980 = income,
         poverty1980 = poverty,
         age1980 = age,
         black1980 = black,
         white1980 = white,
         hispanic1980 = hispanic) %>% 
  select(fips_state_county_code, male1980, female1980, income1980, poverty1980, age1980, black1980, white1980, hispanic1980)

ipums_1990 <- ipums_1980_1990 %>% 
  filter(year == 1990) %>% 
  mutate(male1990 = male,
         female1990 = female,
         income1990 = income,
         poverty1990 = poverty,
         age1990 = age,
         black1990 = black,
         white1990 = white,
         hispanic1990 = hispanic) %>% 
  select(fips_state_county_code, male1990, female1990, income1990, poverty1990, age1990, black1990, white1990, hispanic1990)

rm(ipums_1980_1990)

# Separate `fbi_1980_1990` into individual years
fbi_1980 <- fbi_1980_1990 %>% 
  filter(year == 1980) %>% 
  mutate(population1980 = population,
         arrest_rate1980 = arrest_rate,
         proportion_black1980 = proportion_black,
         proportion_white1980 = proportion_white,
         sale_drug_total_tot_black1980 = sale_drug_total_tot_black,
         sale_drug_total_tot_white1980 = sale_drug_total_tot_white,
         sale_drug_total_tot_arrests1980 = sale_drug_total_tot_arrests) %>% 
  select(fips_state_county_code, population1980, arrest_rate1980, proportion_black1980, proportion_white1980, sale_drug_total_tot_black1980, sale_drug_total_tot_white1980, sale_drug_total_tot_arrests1980)

fbi_1990 <- fbi_1980_1990 %>% 
  filter(year == 1990) %>% 
  mutate(population1990 = population,
         arrest_rate1990 = arrest_rate,
         proportion_black1990 = proportion_black,
         proportion_white1990 = proportion_white,
         sale_drug_total_tot_black1990 = sale_drug_total_tot_black,
         sale_drug_total_tot_white1990 = sale_drug_total_tot_white,
         sale_drug_total_tot_arrests1990 = sale_drug_total_tot_arrests) %>% 
  select(fips_state_county_code, population1990, arrest_rate1990, proportion_black1990, proportion_white1990, sale_drug_total_tot_black1990, sale_drug_total_tot_white1990, sale_drug_total_tot_arrests1990)

panel_1980_1990 <- ipums_1980 

panel_1980_1990 <- panel_1980_1990 %>% 
  left_join(ipums_1990, by = "fips_state_county_code") %>% 
  left_join(fbi_1980, by = "fips_state_county_code") %>% 
  left_join(fbi_1990, by = "fips_state_county_code")
```

Rows contain data on both 1980 and 1990 variables. We may choose to impute missing values, but we have approximately 250 complete observations.

## Return complete panel

```{r}
panel_1980_1990
```

# Analysis

Class 20: https://github.com/THOMASELOVE/432-2022/tree/main/classes/class20


# References

Steven Ruggles, Sarah Flood, Sophia Foster, Ronald Goeken, Jose Pacas, Megan Schouweiler and Matthew Sobek. IPUMS USA: Version 11.0 [dataset]. Minneapolis, MN: IPUMS, 2021. 
https://doi.org/10.18128/D010.V11.0

Kaplan, Jacob. Jacob Kaplan’s Concatenated Files: Uniform Crime Reporting (UCR) Program Data: Arrests by Age, Sex, and Race, 1974-2020: ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds.zip. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2021-09-27. https://doi.org/10.3886/E102263V14-103320

# Session Information

```{r, eval = FALSE}
sessioninfo::session_info()
```
