---
title: "432 Project B"
author: "Jacob James Rich & Steven Mayher"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: show
    self_contained: true
    number_sections: true
    df_print: paged
---

# Setup {-}

```{r setup, message = FALSE}
library(knitr)
library(rmdformats)

## Global options
opts_chunk$set(comment=NA)
opts_knit$set(width=75)
```

## Package Loading {-}

```{r load_packages, message = FALSE}
library(here) #here()
library(doBy) #summaryBy()
library(magrittr)
library(janitor)
library(conflicted) # conflict_prefer()

library(rms) # rcs(), includes Hmisc package
library(MASS) #polr()
library(nnet)
library(skimr)

#install.packages("countreg", repos="http://R-Forge.R-project.org")
library(countreg)

library(broom) # tidy() 
library(car) # outlierTest() 
library(GGally) # ggpairs()
library(mosaic)
library(patchwork)
library(naniar) # n_miss() 
library(ipumsr) # IPUMS data

library(tidyverse)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("sum", "base")
```

## Sources

-Jacob Kaplan's Concatenated Files: Uniform Crime Reporting (UCR) Program [Data](https://www.openicpsr.org/openicpsr/project/102263/version/V14/view?path=/openicpsr/102263/fcr:versions/V14/ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds.zip&type=file): Arrests by Age, Sex, and Race, 1974-2020

- Florida Uniform Crime Report (UCR) [Data](https://www.fdle.state.fl.us/FSAC/CJAB-Home/Uniform-Crime-Report/Data-Archives/UCR-Arrest-Data)

https://www.fdle.state.fl.us/FSAC/Data-Statistics/UCR-Offense-Data.aspx

# Build `FBI_1980_1990`

## 1980 & 1990

```{r}
# Download all files
FBI_1990  <- readRDS(here("ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds", "ucr_arrests_monthly_all_crimes_race_sex_1990.rds")) %>%
  select(year, population, fips_state_county_code, sale_drug_total_tot_white, sale_drug_total_tot_black, sale_drug_total_tot_arrests)

FBI_1980 <- readRDS(here("ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds", "ucr_arrests_monthly_all_crimes_race_sex_1980.rds")) %>%
  select(year, population, fips_state_county_code, sale_drug_total_tot_white, sale_drug_total_tot_black, sale_drug_total_tot_arrests)

FBI_1980_1990 <- rbind(data = FBI_1990, data = FBI_1980)
```

# Build `ipums`

```{r}
ddi <- read_ipums_ddi("usa_00004.xml")

ipums <- read_ipums_micro(ddi)

ipums
```

# Build Panels

## FBI county level

```{r}
# Collapse by year and fips_state_county_code
FBI_county_years <- summaryBy(population + sale_drug_total_tot_white + sale_drug_total_tot_black + sale_drug_total_tot_arrests ~ year + fips_state_county_code, data=FBI_1980_1990, FUN = sum, keep.names = TRUE)

#Determine which counties don't have a match for both years
subset(FBI_1990, !(fips_state_county_code %in% FBI_1980$fips_state_county_code))

# Program "not in" operator
'%!in%' <- function(x,y)!('%in%'(x,y))

# Outcome variable calculations / remove puerto rico, guam, american samoa, virgin islands, florida / remove nonmatched counties
FBI_county_years <- FBI_county_years %>% 
  mutate(arrest_rate = sale_drug_total_tot_arrests / population * 100000,
         proportion_black = sale_drug_total_tot_black / sale_drug_total_tot_arrests,
         proportion_white = sale_drug_total_tot_white / sale_drug_total_tot_arrests,
         fips_state = substr(fips_state_county_code, 1, 2)) %>%
  filter(fips_state_county_code %!in% c(02188, 02100, 02060, 04012, 02060, 02070, 02013, 	04012),
         fips_state %!in% c(72, 66, 60, 78, 12))

#Verify same number of observations for both periods
FBI_county_years %>% 
  filter(year > 1980)

FBI_county_years %>% 
  filter(year < 1990)
```

## FBI national level

```{r}
# Collapse by year only
FBI_years <- summaryBy(population + sale_drug_total_tot_white + sale_drug_total_tot_black + sale_drug_total_tot_arrests ~ year, data=FBI_county_years, FUN = sum, keep.names = TRUE)

# Outcome variable calculations
FBI_years <- FBI_years %>% 
  mutate(arrest_rate = sale_drug_total_tot_arrests / population * 100000,
         proportion_black = sale_drug_total_tot_black / sale_drug_total_tot_arrests,
         proportion_white = sale_drug_total_tot_white / sale_drug_total_tot_arrests)
```

## IPUMS predictor variables

```{r}
# Calculate individual weights
ipums_panel <- ipums %>% 
  mutate(sex_base = if_else(SEX %in% c(1,2), PERWT, 0),
         male = if_else(SEX %in% c(1), PERWT, 0),
         female = if_else(SEX %in% c(2), PERWT, 0),
         income_base = if_else(INCWAGE %!in% c(999999, 999998), PERWT, 0),
         income = if_else(INCWAGE %!in% c(999999, 999998), INCWAGE*PERWT, 0),
         poverty_base = if_else(POVERTY %!in% c(000), PERWT, 0),
         poverty = if_else(POVERTY < 100, if_else(POVERTY == 0, 0, PERWT),0),
         age_base = if_else(AGE < 91, PERWT, 0),
         age = if_else(AGE < 91, AGE*PERWT, 0),
         race_base = if_else(RACE < 8, PERWT, 0),
         black = if_else(RACE %in% c(2), PERWT, 0),
         white = if_else(RACE %in% c(1), PERWT, 0),
         hispanic_base = if_else(HISPAN %!in% c(9), PERWT, 0),
         hispanic = if_else(HISPAN %in% c(1, 2, 3, 4), PERWT, 0)
         )

# Collapse into counties
ipums_panel
```

## Return complete panel

```{r}
FBI_county_years
```

# Analysis

## Plots

```{r}
plot_arrest_rate <- ggplot(FBI_years, aes(x = year)) +
  geom_line(aes(y = arrest_rate), color = "darkred") +
  labs(x = "Year", y = "Arrest Rate per 100,000", title = "Drug Sale Arrest Rate by Year") +
  scale_color_manual(name = "Legend", values = c("Black_Arrest_Rate" = "darkred", "White_Arrest_Rate" = "steelblue", "Total_Arrest_Rate" = "green"))

plot_proportions <- ggplot(FBI_years, aes(x = year)) +
  geom_line(aes(y = proportion_black), color="steelblue") + 
  geom_line(aes(y = proportion_white), color="green") +
  labs(x = "Year", y = "", title = "Proportion by Race") +
  scale_color_manual(name = "Legend", values = c("Proportion_Black" = "darkred", "Proportion_White" = "steelblue", "Total_Arrest_Rate" = "green"))

plot_arrest_rate / plot_proportions
```

# References

Steven Ruggles, Sarah Flood, Sophia Foster, Ronald Goeken, Jose Pacas, Megan Schouweiler and Matthew Sobek. IPUMS USA: Version 11.0 [dataset]. Minneapolis, MN: IPUMS, 2021. 
https://doi.org/10.18128/D010.V11.0

Kaplan, Jacob. Jacob Kaplan’s Concatenated Files: Uniform Crime Reporting (UCR) Program Data: Arrests by Age, Sex, and Race, 1974-2020: ucr_arrests_monthly_all_crimes_race_sex_1974_2020_rds.zip. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2021-09-27. https://doi.org/10.3886/E102263V14-103320

# Session Information

```{r, eval = FALSE}
sessioninfo::session_info()
```
